### 2007-01-09 Documentation/part1.tex
User's manual complete

HIER BIN ICH GERADE [WW].

Example part here for non-standard implementation? [WW]

@@

Using more than six transition parameters in \GLOBES , you will have to maintain the additional parameters. 
For example, there are no standard routines in \GLOBES\ which define projections including more than six parameters, 
which means that you should only use {\tt glbChiNP} further on. In addition, you can still use {\tt glbDefineParams}, but this 
will only access the six standard parameters. You will need to set the additional ones manually using {\tt glbSetOscParams}. 
Do not forget to maintain your non-standard parameters, since any negligence will be punished by un-predicted behavior!

In order to use the non-standard parameters, \GLOBES\ creates any set of oscillation parameters (type {\tt glb\_params})
with the additional parameters. To access the non-standard parameters, you can use the following functions as usual:
\begin{itemize}
\item
{\tt glbSetOscParams}
\item
{\tt glbGetOscParams}
\item
{\tt glbSetProjectionFlag}
\item
{\tt glbGetProjectionFlag}
\end{itemize}
In all these cases, the last parameter {\tt which} can now run from $0$ to {\tt GLB\_OSCP}$-1$. For instance,
\begin{quote}
{\tt
 glbDefineParams(true\_values,0.55,0.15,0.78,0.0,0.000082,0.0022); \\
 glbSetOscParams(true\_values,0.0,6);
}
\end{quote}
sets all of the oscillation parameters if {\tt GLB\_OSCP}=7. In addition, the function {\tt glbCopyParams}
can be used to copy all parameters inluding the non-standard ones. For projections,
do not use any of the {\tt glbChi...} functions anymore except from {\tt glbChiNP} in order to have a
predictive behavior of the extra dimenstions. Similar to the oscillation parameters, you can define
the marginalization over the extra parameter(s) by {\tt glbSetProjectionFlag}. You can find a simple
example using non-standard physics on page~\pageref{ex:nsphysics}.

\example{Non-standard process on Hamiltonian level}{
\label{ex:nsphysics}
\index{norm}{Hamiltonian}
Let us discuss an example of a very simple Hamiltonian-level non-standard flavor-changing effect which 
couples $\nu_\mu$ to $\nu_\tau$ in flavor base. This effect can be parameterized
relative to the standard matter effect acting on $\nu_e$ by an additional
parameter, which we adopt as our 7th oscillation parameter. In order
to incorporate this effect, we change a few lines in the function {\tt ComputeMixingMatrix}
in {\tt glb\_probability.cc}:
\begin{quote}
{\tt {\footnotesize
double msw = 0.1*1.5*anti\_neutrinos*(10E-4)*neutrino\_energy*0.5*matter\_density; \\
md[0][0] += msw; \\
md[1][2] += msw*nsparams[0]; md[2][1] += msw*nsparams[0];
}}
\end{quote}
Here we have assume a real contribution which has to be added in a Hermitian way (the Hamiltonian
has to be Hermitian), where the first line is the standard matter effect acting on $\nu_e$ only 
($2 \sqrt{2} G_F n_e E$ expressed in our units).

In the application software, we can now use this additional parameter. Let us assume that we
want to compute the projection onto the $\stheta$-axis when we do not know the size of the  non-standard
effect, \ie, we keep it completely free but assume it to be zero as null hypothesis:
\begin{quote}
{\tt {\footnotesize
 \#define MY\_FV 6 \\
 ... \\
 glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm); \\
 glbSetOscParams(true\_values,0.0,MY\_FV); \\
 glbDefineParams(input\_errors,0.1*theta12,0,0,0,0.1*sdm,0); \\
 glbSetOscParams(input\_errors,0,MY\_FV);  \\
 glbSetDensityParams(input\_errors,0.05,GLB\_ALL); \\
 glbSetOscillationParameters(); glbSetRates(); \\
 glbSetCentralValues(true\_values);glbSetInputErrors(input\_errors); \\
 \\
 glbDefineProjection(my\_proj,GLB\_FREE,GLB\_FIXED,GLB\_FREE,\\
  \hspace*{2cm} GLB\_FREE,GLB\_FREE,GLB\_FREE); \\
 glbSetProjectionFlag(my\_proj,GLB\_FREE,MY\_FV); \\
 glbSetDensityProjectionFlag(my\_proj,GLB\_FREE,GLB\_ALL); \\
 glbSetProjection(my\_proj);
 \\
 glbCopyParams(true\_values,fit\_values); \\
 glbSetOscParams(fit\_values,0.1,GLB\_THETA\_13); \\
 res=glbChiNP(fit\_values,minimum,GLB\_ALL); 
}}
\end{quote}
This code fragment illustrates the implementation of a {\tt glbChiTheta13} function
including the non-standard effect, which is being marginalized over.
Alternatively, one could define a {\tt glbChiTheta13} fixing the non-standard
parameter, or include external constrainys to limit its size. Therefore, 
there is no ``standard'' treatment of new physics effects incorporated
in \GLOBES .
}

>>>>

Let us now illustrate the implementation of non-standard physics with a simple example, which can
be found as {\tt example6.c} in the example directory.
This example is a simplified version of \Sec~4 of 
\Ref~\cite{Blennow:2005yk}.
It uses an analytical probability calculation for a reactor experiment with a baseline being treated
in vacuum for simplicity. The non-standard effect is the loss of coherence because of wave packet decoherence
or any other such effect, \ie, one more oscillation parameter. We need to define two functions to access a set of 
global oscillation parameters
double th12, th13, th23, deltacp, sdm, ldm, sigma\_E; \\
int my\_set\_oscillation\_parameters(glb\_params p, void *user\_data) \\
 \hspace*{0.5cm} th12    = glbGetOscParams(p, GLB\_THETA\_12); \\
 \hspace*{0.5cm} th13    = glbGetOscParams(p, GLB\_THETA\_13); \\
 \hspace*{0.5cm} th23    = glbGetOscParams(p, GLB\_THETA\_23); \\
 \hspace*{0.5cm} deltacp = glbGetOscParams(p, GLB\_DELTA\_CP); \\
 \hspace*{0.5cm} sdm     = glbGetOscParams(p, GLB\_DM\_21) * 1.0e-18;  \\
 \hspace*{0.5cm} ldm     = glbGetOscParams(p, GLB\_DM\_31) * 1.0e-18;  \\
 \hspace*{0.5cm} sigma\_E = glbGetOscParams(p, GLB\_SIGMA\_E); \\
  return 0; \\
int my\_get\_oscillation\_parameters(glb\_params p, void *user\_data) \\
  \hspace*{0.5cm} glbSetOscParams(p, th12, GLB\_THETA\_12); \\
  \hspace*{0.5cm} glbSetOscParams(p, th13, GLB\_THETA\_13); \\
  \hspace*{0.5cm} glbSetOscParams(p, th23, GLB\_THETA\_23); \\
  \hspace*{0.5cm} glbSetOscParams(p, deltacp, GLB\_DELTA\_CP); \\
  \hspace*{0.5cm} glbSetOscParams(p, sdm*1.0e18, GLB\_DM\_21);   \\
  \hspace*{0.5cm} glbSetOscParams(p, ldm*1.0e18, GLB\_DM\_31);   \\
  \hspace*{0.5cm} glbSetOscParams(p, sigma\_E, GLB\_SIGMA\_E); \\
  \hspace*{0.5cm} return 0; \\
\} 
Furthermore, we need to define the probability matrix calculation itself -- in this case,
we perform an analytical calculation:
int my\_probability\_matrix(double P[3][3], int cp\_sign, double E, int psteps,
                          const double *length, const double *density,
                          double filter\_sigma, void *user\_data) \\
 \hspace*{0.5cm}  int i, j; \\
 \hspace*{0.5cm} double L, Delta21, Delta31, Delta32; \\
 \hspace*{0.5cm} double D21, D31, D32, s13, c13, s12, c12, t; \\
  \hspace*{0.5cm} /* Set all probabilities to zero initially */ \\
  \hspace*{0.5cm} for (i=0; i < 3; i++) for (j=0; j < 3; j++) P[i][j] = 0.0; \\
 \hspace*{0.5cm} /* Calculate total baseline */ \\
 \hspace*{0.5cm} L = 0.0; for (i=0; i < psteps; i++) L += length[i]; \\
 \hspace*{0.5cm} L = KM\_TO\_EV(L) * 1.0e9;      /* Convert to GeV$^{-1}$ */ \\
 \hspace*{0.5cm} /* Compute P\_ee */ \\
  \hspace*{0.5cm} s12 = sin(th12); c12 = cos(th12); \\
  \hspace*{0.5cm} s13 = sin(th13); c13 = cos(th13); \\
  \hspace*{0.5cm} t = L / (4.0 * E); \\ 
  \hspace*{0.5cm} Delta21 = sdm * t; Delta31 = ldm * t; \\
  \hspace*{0.5cm} Delta32 = Delta31 - Delta21; \\
  \hspace*{0.5cm} t = M\_SQRT2 * sigma\_E / E; \\
  \hspace*{0.5cm} D21 = exp(-square( Delta21 * t )); \\
  \hspace*{0.5cm} D31 = exp(-square( Delta31 * t )); \\
  \hspace*{0.5cm} D32 = exp(-square( Delta32 * t )); \\
  \hspace*{0.5cm} P[0][0] = square(square(c13)) * ( 1 - 2.0*square(s12*c12)* \\
  \hspace*{1cm} (1 - D21*cos(2.0*Delta21)) ) 
  \hspace*{1cm} + 2.0*square(s13*c13) * \\
  \hspace*{1cm} ( D31*square(c12)*cos(2.0*Delta31) \\
  \hspace*{1cm} + D32*square(s12)*cos(2.0*Delta32) )  + square(square(s13)); \\
  \hspace*{0.5cm} return 0; \\
\}
Now we can register the probability engine after {\tt glbInit}:
  glbInit(argv[0]); \\
  glbRegisterProbabilityEngine(7,      /* Number of parameters */ \\
  \hspace*{1cm}                             \&my\_probability\_matrix, \\
  \hspace*{1cm}                             \&my\_set\_oscillation\_parameters, \\
  \hspace*{1cm}                             \&my\_get\_oscillation\_parameters, \\
  \hspace*{1cm}                             NULL); 
We will demonstrate in the next section how to access the non-standard physics 
parameter in the application software.

\example{Decoherence in neutrino oscillations}{
\label{ex:nsphysics}
\index{norm}{Decoherence}
The following code fragment from {\tt example6.c} calculates 
a fit region on $\stheta$-$\sigma_E$-space, where $\sigma_E$ is the
non-standard decoherence parameter (\cf, \Ref~\cite{Blennow:2005yk}):
\begin{quote}
{\tt {\footnotesize
  const GLB\_SIGMA\_E = 6; \\ \\
  /* Fix theta13 and sigma\_E, and marginalize over all other parameters except deltaCP and th23, which do not enter P\_ee */ \\
  glbDefineProjection(myproj, GLB\_FREE, GLB\_FIXED, GLB\_FIXED,
    GLB\_FIXED, \\ \hspace*{0.5cm} GLB\_FREE, GLB\_FREE); \\
  glbSetDensityProjectionFlag(myproj, GLB\_FIXED, GLB\_ALL); \\
  glbSetProjectionFlag(myproj, GLB\_FIXED, GLB\_SIGMA\_E);  \\
  glbSetProjection(myproj); \\ 
\\
  for(x=0; x < 0.05+0.001; x+=0.005)      /* th13 loop */ \\
  for(y=0.0; y < 0.010+0.001; y+=0.001)   /* sigma\_E loop */ \\
    \{ \\
   \hspace*{0.5cm}    /* Set vector of test=fit values */ \\
   \hspace*{0.5cm}    thetheta13=asin(sqrt(x))/2.0; \\
   \hspace*{0.5cm}    glbSetOscParams(test\_values,thetheta13,GLB\_THETA\_13); \\
   \hspace*{0.5cm}    glbSetOscParams(test\_values,y,GLB\_SIGMA\_E); \\
     \\
   \hspace*{0.5cm}    /* Compute Chi2 with correlations */ \\
   \hspace*{0.5cm}    res=glbChiNP(test\_values,NULL,GLB\_ALL); \\
  \\
   \hspace*{0.5cm}    AddToOutput(x,y,res); \\
    \} 
}}
\end{quote}

The result is represented by the dark curves
 in the following figure (similar to \Ref~\cite{Blennow:2005yk}):
\begin{center}
\colorbox{white}{\includegraphics[width=6cm]{reactorNSI}}
\end{center}
}

@@

Using more than six parameters in \GLOBES , you will have to maintain the additional parameters. 
For example, there are no standard routines in \GLOBES\ which define projections including more than six parameters, 
which means that you should only use {\tt glbChiNP} further on. In addition, you can still use {\tt glbDefineParams}, but this 
will only access the six standard parameters. You will need to set the additional ones manually using {\tt glbSetOscParams}. 
Do not forget to maintain your non-standard parameters, since any negligence will be punished by un-predicted behavior!

In order to use the non-standard parameters, \GLOBES\ creates any set of oscillation parameters (type {\tt glb\_params})
with the additional parameters. To access the non-standard parameters, you can use the following functions as usual:
\begin{itemize}
\item
\GLB{SetOscParams}
\item
\GLB{GetOscParams}
\item
\GLB{SetProjectionFlag}
\item
\GLB{GetProjectionFlag}
\end{itemize}
In all these cases, the last parameter {\tt which} can now run from $0$ to \GLB{GetNumOfOscParams}{\tt ()-1}. For instance,
\begin{quote}
{\tt
 glbDefineParams(true\_values,0.55,0.15,0.78,0.0,0.000082,0.0022); \\
 glbSetDensityParams(true\_values,1.0,GLB\_ALL); \\
 glbSetOscParams(true\_values,0.0,6);
}
\end{quote}
sets all of the oscillation parameters if you have one additional parameter. 
In addition, the function {\tt glbCopyParams}
can be used to copy all parameters inluding the non-standard ones. For projections,
do not use any of the {\tt glbChi...} functions anymore except from {\tt glbChiNP} in order to have a
predictive behavior of the extra dimenstions. Similar to the oscillation parameters, you can define
the marginalization over the extra parameter(s) by {\tt glbSetProjectionFlag}. You can find a simple
example using non-standard physics with the code from the last section on page~\pageref{ex:nsphysics}.

\chapter{Experimental features}

Here we describe experimental features currently being implemented in \GLOBES . These features have not yet been tested extensively and should be used with care. They may evolve into standard features in the future, or they may not be supported anymore at a certain point. In general, the number of experiment features is small in a new release version, and increases towards a new version number.

\subsection*{\GLOBES\ 3.0 and higher}

One can change the minimization algorithm in \GLOBES\ with
\begin{function}
\GLBNS{SelectMinimizer}
{\tt int glbSelectMinimizer(int minimizer\_ID)} selects the minimizer {\tt GLB\_MIN\_NESTED\_POWELL}, {\tt GLB\_MIN\_DEFAULT}, or the hybrid minimzer {\tt GLB\_MIN\_POWELL} by the parameter {\tt minimizer\_ID}. 
{\tt GLB\_MIN\_NESTED\_POWELL} is the currently implemented standard algorithm, {\tt GLB\_MIN\_DEFAULT} chooses the standard algorithm at any given time, and {\tt GLB\_MIN\_POWELL} is a faster, yet rarely tested hybrid
minimizer.
\end{function}
Compared to the standard minimization which is performed for systematics first and then for the
oscillation parameters, the hybrid minimizer mixes the systematics and oscillation parameter minimizations. This method is much faster, but new (unknown) correlations between systematics parameters and oscillation parameters may emerge. Since one can change the minimizer at any time, it is recommended that one cross check the minimization for a particular experiment.
